---
title: "RMarkdown"
subtitle: "Show and reproduce with RMarkdown"
author: "Bea Hernández"
date: "`r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

# RMarkdown

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
class: middle

- R is **‘GNU S’** (free software under the GPL License)<sup>1</sup>
- R is **language and environment**
- R is for **statistical computing** and **graphics**

.footnote[
[1] https://cran.r-project.org/
]

---
class: inverse, center, middle

# Get Started

---

# Install R

1. Download from [R-project](https://www.r-project.org/)
2. Check your installation: in the terminal type `R`

```bash
$ R

R version 3.5.0 (2018-04-23) -- "Joy in Playing"
Copyright (C) 2018 The R Foundation for Statistical Computing
Platform: x86_64-apple-darwin17.5.0 (64-bit)
```

--

You are recommended to use the [RStudio IDE](https://www.rstudio.com/products/rstudio/), but you do not have to. There are other IDEs:

- [Emacs](https://ess.r-project.org/)
- [PyCharm](https://plugins.jetbrains.com/plugin/6632-r-language-support)

--

Or use the terminal:

--

```bash
$ Rscript my_script.R
```

---
class: center, middle, inverse

# R basics

---

# Basic math

```{r}
1 + 2 + 3
1 + 2 * 3
```

--

The power of parenthesis
```{r}
(1 + 2) * 3
```

--

Other computations

```{r eval = F}
2 ^ 10
10 / 2
```

---
# Comparisons

```{r}
3 == 4

FALSE & TRUE

FALSE | TRUE
```

--

Use `&` and `|` in logical expressions involving the logical values TRUE and FALSE. 

Use `&&` and `||` for the flow-of-control expressions inside if and while statements (short-circuit evaluation).

---
# Vectors

`c` for *combine*

```{r}
v <- c(0, 1, 1, 2, 3, 5, 8)
v
```
--
```{r}
w <- 1:10
w
```
--
```{r}
c(v, w)
```

---
# Vectors

```{r}
c(1, 2, 3, 4) * c(10, 20, 30, 40)
```
--
```{r}
c(1, 2, 3, 4) + 1
```
--
```{r}
1 / c(1, 2, 3, 4, 5)
```
--
```{r}
c(1, 2, 3, 4, 5) + c(10, 100)
```
---
# Variables and functions

- Assign with `<-`
- Call a function with `f(argument1, argument2, ...)`

--

```{r}
exp(1)
```
--
```{r}
b <- c(1,2,3,4,5,6,7,8,9,10,11,12)
b[1:3]
```
--
```{r}
3 -> three
three
```
--
```{r}
f <- function(x,y) {c(x+1, y+1)}
f
```


---
class: middle, inverse, center
# Data Structures

---
class: middle

- An array is a multidimensional vector. Vectors and arrays are stored the same way internally, but an array may be displayed differently and accessed differently.
- Data frame has column names and row names. Data frame let you treat some features as factor.

---
# Lists

```{r}
e <- list(thing="hat")
e
```
--
```{r}
g <- list("this list references another list", e)
g
```

If the variable lst contains a list, it can be indexed in two ways: `lst[[n]]` is the nth element of the list; whereas `lst[n]` is a list whose only element is the nth element of lst. 

---
# Array and Matrix

```{r}
a <- matrix(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12), nrow = 3, ncol = 4)
a
```

--

```{r}
w <- array(c(1,2,3,4,5,6,7,8,9,10,11,12,13,14,15,16,17,18),dim=c(3,3,2))
w[, , 1]
```

---
# Data Frames

```{r}
arrest <- USArrests
head(arrest)
names(arrest)
row.names(arrest)[1:5]
```

---
# Special Values

```{r}
NA
Inf
-Inf
NaN
NULL
```
---
class: middle, inverse, center
# Functions and packages

---
# Functions

---
# Packages
---
class: middle, inverse, center

# HALP!

---
class: middle, center

- [CRAN Packages](https://cran.r-project.org/)
- `Help`
- Vignettes
- `?`

---

class: center, middle

Exercises

**exercises/r_lang.R**