---
title: "Tidyverse"
subtitle: "Tidyverse for data "
author: "Bea Hernández"
date: "`r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

# Tidyverse

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
# What is the Tidyverse?

[Tidyverse](https://www.tidyverse.org/) is a bundle of packages which can be installed separated or together with `tidyverse`, and they share an underlying design philosophy, grammar, and data structure.

--

The main packages of this bundle are:

- **ggplot2** is a system for declaratively creating graphics, based on The Grammar of Graphics
- **dplyr** provides a grammar of data manipulation
- **tidyr** provides a set of functions that help you get to [tidy data](http://vita.had.co.nz/papers/tidy-data.pdf)
- **readr** provides a fast and friendly way to read rectangular data
- **purrr** enhances R's functional programming toolkit
- **tibble** is a modern re-imagining of the data frame
- **stringr** provides a cohesive set of functions designed to make working with strings as easy possible
- **forcats** provides a suite of useful tools that solve common problems with factors

---
# Why I use (and teach) the tidyverse?

It has huge advantages:

- All packages follow the same grammar
- It's easier to perform data management compared to base R
- Readabily and reproducibility are improved

--

## But, isn't it slower than base R?

*Sometimes*. Even so, base R is not the fastest way to do data science in R either. We can work with `data.table`, `h2o`, `Rcpp`, etc, to manage data and improve our speed in R. 

Anyway, this will always depend on many variables, the bottom line is:

**The Exploratory Data Analysis is way easier to do in the tidyverse**

If you need to optimize your workflow (speedwise) then you'll need to do some researching.

---
class: inverse, center, middle

# Let's work

---
# Do and learn

Install and load the `tidyverse` bundle package:

--

```{r eval=F}
install.packages('tidyverse')
library(tidyverse)
```

```{r include=F}
library(tidyverse)
library(nycflights13)
```
--

Install and load the `nycflights13` package:

--

```{r eval=F}
install.packages('nycflights13')
library(nycflights13)
```

---
# Get basic information

*Use the `str` and `summary` on the flights dataset to answer the following questions about the dataset*:

1. What are the classes of the dataset?
2. What is the class of the column *tailnum*?
3. What is the min distance of the flights?
4. What if the median of the datetime?

---
# Filter with dplyr::filter

Use the function `filter` to find all flights that:

- Had an **arrival delay** of **two or more** hours. Hint: `>=`

--

```{r eval=F}
filter(flights, arr_delay >= 120)
```

--

- Flew to Houston (IAH or HOU). Hint: `|`

--

```{r eval=F}
filter(flights, dest == "IAH" | dest == "HOU")
```

--

- Were operated by United (*UA*), American (*AA*), or Delta (*DL*). Hint: `%in%`

--

```{r eval=F}
airlines
filter(flights, carrier %in% c("AA", "DL", "UA"))
```

--

- Arrived more than two hours late, but didn’t leave late

--

```{r eval=F}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

---
# Filter with dplyr::filter

- Departed in summer (July, August, and September). Hint: `between`

--

```{r eval=F}
filter(flights, between(month, 7, 9))
```

---
# Arrange rows with dplyr::arrange

Use the function `arrange` to set the following order:

- Sort all missing values to the start. Hint: `is.na()`

--

```{r, eval = F}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

--

- Sort flights to find the most delayed flights. 

--

```{r, eval = F}
arrange(flights, desc(dep_delay))
```

--

- Which flights traveled the shortest?

--

```{r, eval = F}
arrange(flights, distance)
```

---
# Select columns with dplyr::select

Use the function `select` to select the columns:

- Brainstorm as many ways as possible to select *dep_time*, *dep_delay*, *arr_time*, and *arr_delay* from flights.

--

```{r, eval = F}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

--


```{r, eval = F}
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")
```

--


```{r, eval = F}
select(flights, 4, 6, 7, 9)
```

--

```{r, eval = F}
select(flights, one_of(c("dep_time", "dep_delay", "arr_time", "arr_delay")))
```

--


```{r, eval = F}
select(flights, starts_with("dep_"), starts_with("arr_"))
```

---
# Add new variables with dyplr::mutate

Use `mutate` to add new computed columns:

- Compare air_time with arr_time - dep_time. What do you expect to see? What do you see? What do you need to do to fix it?

--

*I expect that `air_time` is the difference between the arrival (`arr_time`) and departure times (`dep_time`). In other words, `air_time = arr_time - dep_time`*

```{r, eval=F}
flights_airtime <-
  mutate(flights,
    dep_time = (dep_time %/% 100 * 60 + dep_time %% 100) %% 1440,
    arr_time = (arr_time %/% 100 * 60 + arr_time %% 100) %% 1440,
    air_time_diff = air_time - arr_time + dep_time
  )

filter(flights_airtime, air_time_diff != 0)
```

---
# Summarise data with dplyr::summarise

Use `group_by` and `summarise` to get insights about the dataset:

- Look at the number of cancelled flights per day. Is there a pattern? Is the proportion of cancelled flights related to the average delay? 

We are going to introduce the *pipe* (`%>%`) and `ggplot` to see our result. The pipe prevents us to create a lot of intermediate variables, and occupy necessary memory. Following the previous question:

--

```{r eval=F}
cancelled_delayed <- flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay)))

```

Which means: 

- Over the dataset flights, I create a new column cancelled.
- I can skip the *data* parameter because the pipe points to it.
- I can perform multiple operations and assign them to *cancelled_delayed*.

---
# Summarise data with dplyr::summarise

Group the data:

```{r eval=F}
cancelled_delayed <-
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) 
```

--

Summarise the data:

```{r echo = T, results = 'hide'}
cancelled_delayed <-
  flights %>%
  mutate(cancelled = (is.na(arr_delay) | is.na(dep_delay))) %>%
  group_by(year, month, day) %>%
  summarise(
    prop_cancelled = mean(cancelled),
    avg_dep_delay = mean(dep_delay, na.rm = TRUE)
  )
```

**The pipe allows to do function composition in a readable way.**

---
# Plot the summary with ggplot2::ggplot

`ggplot` works by layers, it applies the same philosophy as dplyr:

--

```{r fig.height=2, fig.align='center'}
ggplot(cancelled_delayed, aes(x = avg_dep_delay, prop_cancelled))
```

--

```{r fig.height=2, fig.align='center'}
ggplot(cancelled_delayed, aes(x = avg_dep_delay, prop_cancelled)) +
  geom_point()
```

---
# Plot the summary with ggplot2::ggplot

```{r, message=F, fig.height=5, fig.align='center'}
ggplot(cancelled_delayed, aes(x = avg_dep_delay, prop_cancelled)) +
  geom_point() +
  geom_smooth()
```