---
title: "Tidyverse"
subtitle: "Tidyverse VS. Base"
author: "Bea Hernández"
date: "`r Sys.Date()`"
output: 
  xaringan::moon_reader:
    css: [default, metropolis, metropolis-fonts]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: inverse, center, middle

# Tidyverse

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

---
# What is the Tidyverse?

[Tidyverse](https://www.tidyverse.org/) is a bundle of packages which can be installed separated or together with `tidyverse`, and they share an underlying design philosophy, grammar, and data structure.

--

The main packages of this bundle are:

- **ggplot2** is a system for declaratively creating graphics, based on The Grammar of Graphics
- **dplyr** provides a grammar of data manipulation
- **tidyr** provides a set of functions that help you get to [tidy data](http://vita.had.co.nz/papers/tidy-data.pdf)
- **readr** provides a fast and friendly way to read rectangular data
- **purrr** enhances R's functional programming toolkit
- **tibble** is a modern re-imagining of the data frame
- **stringr** provides a cohesive set of functions designed to make working with strings as easy possible
- **forcats** provides a suite of useful tools that solve common problems with factors

---
# Why I use (and teach) the tidyverse?

It has huge advantages:

- Each package follows the same grammar
- It's easier to perform data management compared to base R
- Readabily is improved

--

## But, isn't it slower?

Sometimes. Even so, base R is not the fastest way to do data science in R either. We can work with `data.table`, `h2o`, `Rcpp`, etc, to manage data and improve our speed in R. 

Anyway, this will always depend on many variables, the bottom line is:

**The data wrangling step of ML is way easier to do in the tidyverse**

If you need to optimize your workflow (speedwise) then you'll need to do some researching.

---
class: inverse, center, middle

# Let's work

---
# Do and learn

Install and load the `tidyverse` bundle package:

--

```{r eval=F}
install.packages('tidyverse')
library(tidyverse)
```

--

Install and load the `nycflights13` package:

--

```{r eval=F}
install.packages('nycflights13')
library(nycflights13)
```

---
# Get basic information

Use the `str` and `summary` on the flights dataset to answer the following questions:

1. What are the classes of the dataset?
2. What is the class of the column *tailnum*?
3. What is the min distance of the flights?
4. What if the median of the datetime?

---
# Filter with dplyr::filter

Use the function `filter` to find all flights that:

--

- Had an **arrival delay** of **two or more** hours. Hint: `>=`

--

```{r eval=F}
filter(flights, arr_delay >= 120)
```

--

- Flew to Houston (IAH or HOU). Hint: `|`

--

```{r eval=F}
filter(flights, dest == "IAH" | dest == "HOU")
```

--

- Were operated by United, American, or Delta. Hint: `%in%`

--

```{r eval=F}
airlines
filter(flights, carrier %in% c("AA", "DL", "UA"))
```

--

- Arrived more than two hours late, but didn’t leave late

--

```{r eval=F}
filter(flights, arr_delay > 120, dep_delay <= 0)
```

---
# Filter with dplyr::filter

- Departed in summer (July, August, and September). Hint: `between`

--

```{r eval=F}
filter(flights, between(month, 7, 9))
```

---
# Arrange rows with dplyr::arrange

Use the function `arrange` to set the following order:

- Sort all missing values to the start. Hint: `is.na()`

--

```{r, eval = F}
arrange(flights, desc(is.na(dep_time)), dep_time)
```

--

- Sort flights to find the most delayed flights. 

--

```{r, eval = F}
arrange(flights, desc(dep_delay))
```

--

- Which flights traveled the shortest?

--

```{r, eval = F}
arrange(flights, distance)
```

---
# Select columns with dplyr::select

Use the function `select` to select the columns:

- Brainstorm as many ways as possible to select dep_time, dep_delay, arr_time, and arr_delay from flights.

--

```{r, eval = F}
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

--


```{r, eval = F}
select(flights, "dep_time", "dep_delay", "arr_time", "arr_delay")
```

--


```{r, eval = F}
select(flights, 4, 6, 7, 9)
```

--

```{r, eval = F}
select(flights, one_of(c("dep_time", "dep_delay", "arr_time", "arr_delay")))
```

--


```{r, eval = F}
select(flights, starts_with("dep_"), starts_with("arr_"))
```

---
# Add new variables with dyplr::mutate

